### 2021-12-13 input/proposal.tex
meow

Algorithm~\ref{algo:add} shows the few instructions that implement a
\emph{reactive} forwarding of best to reach consistent partitioning in
a static network where \processes never join, crash, nor leave the
system. As soon as a \process receives a better message, it delivers
and forwards it. The trade-off consists in decreasing termination time
at cost of increased traffic. \TODO{For the sake of simplicity and to
  remove obvious loops, a \process does not forward a notification to
  its sender (Line~\ref{line:accumulator})}. Figure~\ref{fig:add}
illustrates the behavior of this algorithm on a system comprising 4
\processes $a$, $b$, $c$, and $d$. Both $a$ and $d$ become sources.
They \NAMEB notifications $\alpha_a^0$ and $\alpha_d^0$. They
initialize their own state with the lowest value $0$ (see
Line~\ref{line:lowestbound}), and send a message to each of their
neighbors by accumulating the corresponding edge weight (see
Line~\ref{line:accumulator}). In Figure~\ref{fig:addC}, $b$ receives
$\alpha_{d}^{1}$. Since it improves its own partition distance, it
keeps it and forwards it to its neighbors. In Figure~\ref{fig:addD},
$b$ discards $\alpha_{a}^{2}$, for it does not improve its partition
distance. Neither $c$ nor $d$ will ever acknowledge that Source~$a$
exists. Ultimately, \processes discard last transiting messages.
% Despite the obvious lack of traffic optimization,
The system reaches consistent partitioning.

\noindent \TODO{In terms of space complexity.} In terms of number of
messages, in the average case, a \process $x$ chosen uniformly at
random among all \processes creates a logical partition. Its messages
$\alpha_x$ spread through the network until reaching \processes that
belong to another partition closer to them. This splits partitions in
half on average. Overall, the $n^{th}$ new partition comprises
\smash{$\mathcal{O}(\frac{|V|}{2^{\lfloor \log_2 n \rfloor}})$}
\processes. This decreases every new partition until reaching one
\process per new partition: its source. The average number of messages
per \process is
\smash{$\mathcal{O}(\frac{\overline{|O|}}{2^{\lfloor \log_2 n
      \rfloor}})$}, where \smash{$\overline{|O|}$} is the average
number of neighbors per \process.

While only adding logical partitions to the distributed system is
straightforward, removing them introduces additional complexity caused
by concurrent operations.

@@

Delete operations bring a new notion of order between events, and most
importantly between message deliveries. A \process $s$ that performs a
\texttt{Del} operation implicitly states that all its preceding
\texttt{Add} operations become obsolete. As consequence, their results
-- scoped broadcast in the form of \texttt{\underline{A}dd}
notifications $\alpha_s$ -- should be canceled by the corresponding
\texttt{\underline{D}el} notifications $\delta_s$. \TODO{From algo, we
  see that last is also best, so we want to exploit this. but why ? or
  should we show ``forwarding all staleness'' first.}

@@

\begin{definition}[\label{def:lwo}Stale messages]
  Only the latest broadcast of a \node matters.  A message $m$ conveys
  \underline{s}tale control information $\mathcal{S}(m)$ as soon as
  its broadcaster broadcasts another message: $\mathcal{S}(\alpha_x) =
  \exists \delta_x: b_x(\alpha_x) \rightarrow b_x(\delta_x)$.  A
  \process only delivers or sends messages that it assumes
  \underline{u}p-\underline{t}o-\underline{d}ate.
  % : $s_x(\alpha_s)
  % \implies \delta_s \not\in R_x$.
  For convenience, we define \TODO{$\textup{minUTD(}D\textup{)} = \min
  \{\alpha_s^d | \alpha_s^d \in D \wedge \delta_s \notin D\}$}, and
  $\textup{last(}D\textup{)} = \alpha_s \in D: \nexists \alpha_{s'}
  \in D: d_x(\alpha_s) \rightarrow d_x(\alpha_{s'})$.
  %  \TODO{Only keep last?  Tothink what happens if we keep all as in
%    def now.}
\end{definition}

A naive attempt at implementing DCP resembles echoes in
acoustics. Following the principles of scoped broadcast as stated in
Definition~\ref{def:scoped}, a \process that receives a staleness
notification forwards it only if the latter targets its current
partition. The staleness notifications propagate through the network
as long as they remain in the deleted partition. When they reach the
bordering \processes of the deleted partition, it triggers a
competition -- or an echo -- that goes backward to fill the gap left
open by removals. \TODO{as in FB.}

@@

\begin{definition}[\label{def:fs}\underline{F}orwarding of
  \underline{s}taleness (FS)]
  % 
  \TODO{A source broadcasts.}
  A \process $x$ delivers and forwards \TODO{at most once} a received staleness
  notification if it targets its best up-to-date delivered message:\\
  \TODO{$(\delta_s \in R_x \wedge \alpha_s = \textup{minUTD}~D_x) \iff
  \eventually (d_x(\delta_s) \rightarrow f_x(\delta_s))$.}\\
  $\textup{last}~D = \alpha_s \wedge d_x(\alpha_s) \rightarrow r_x(\delta_s)
  \implies d_x(\delta_s) \rightarrow f_x(\delta_s)$.
\end{definition}

\begin{definition}[\label{def:fbplus}$\textup{FB}^+$: echos] FB where
  a \process $x$ that receives but does not deliver a staleness
  notification sends back --~or echoes~-- its best up-to-date
  delivered message:
  $\textup{last}~D_x = \alpha_s^d \wedge \delta_s
  \not\in D_x \wedge d_x(\alpha_s^d)\rightarrow r_{xy}(\delta_{s'}) \implies
  s_{xy}(\alpha_s^d\oplus^{w_{xy}})$.

  % $r_{xy}(\alpha_s)
  % \wedge \delta_s \in D_x \implies s_{xy}(\alpha_{s'}^d\oplus
  % ^{w_{xy}})$ with $\alpha_{s'^d} = \min D_x \wedge \delta_{s'} \not\in
  % D_x$. \TODO{Rework min}

\end{definition}

@@

\begin{definition}[\label{def:fsplus}$\textup{FS}^+$: forwarding
  staleness downstream]
  \TODO{Any process broadcast.}  A node $x$ delivers and forwards a
  received staleness notification if it is not yet delivered and it
  comes from the path of delivery of its best up-to-date delivered
  message. \TODO{more.}
\end{definition}

>>>>

Algorithm~\ref{algo:add} shows the few instructions that implement a
\emph{reactive} forwarding of best to reach consistent partitioning in
a static network where \processes never join, crash, nor leave the
system. As soon as a \process receives a better message, it delivers
and forwards it. The trade-off consists in decreasing termination time
at cost of increased traffic.  Figure~\ref{fig:add} illustrates the
behavior of this algorithm on a system comprising 4 \processes $a$,
$b$, $c$, and $d$. Both $a$ and $d$ become sources.  They \NAMEB
notifications $\alpha_a^0$ and $\alpha_d^0$. They initialize their own
state with the lowest value $0$ (see Line~\ref{line:lowestbound}), and
send a message to each of their neighbors by accumulating the
corresponding edge weight (see Line~\ref{line:accumulator}). In
Figure~\ref{fig:addC}, $b$ receives $\alpha_{d}^{1}$. Since it
improves its own partition distance, it keeps it and forwards it to
its neighbors. In Figure~\ref{fig:addD}, $b$ discards
$\alpha_{a}^{2}$, for it does not improve its partition
distance. Neither $c$ nor $d$ will ever acknowledge that Source~$a$
exists. Ultimately, \processes discard last transiting messages.
Despite the obvious lack of traffic optimization, the system reaches
consistent partitioning.

Each \process has a space complexity of $\mathcal{O}(\log_2(d \cdot
|V|))$, and generates messages of identical space complexity. In terms
of number messages, in the average case, a \process $x$ chosen
uniformly at random among all \processes creates a logical
partition. Its messages $\alpha_x$ spread through the network until
reaching \processes that belong to another partition closer to
them. This splits partitions in half on average. Overall, the $n^{th}$
new partition comprises \smash{$\mathcal{O}(\frac{|V|}{2^{\lfloor
      \log_2 n \rfloor}})$} \processes. This decreases every new
partition until reaching one \process per new partition: its
source. The average number of messages per \process is
\smash{$\mathcal{O}(\frac{\overline{|O|}}{2^{\lfloor \log_2 n
      \rfloor}})$}, where \smash{$\overline{|O|}$} is the average
number of neighbors per \process.

While only adding logical partitions to the distributed system is
straightforward and lightweight, removing them introduces additional
complexity and costs caused by concurrent operations.

@@

Delete operations bring a new notion of order between events, and most
importantly between message deliveries. A \process $s$ that performs a
\texttt{Del} operation implicitly states that all its preceding
\texttt{Add} operations become obsolete. As consequence, their results
-- scoped broadcast in the form of \texttt{\underline{A}dd}
notifications $\alpha_s$ -- should be canceled by the corresponding
\texttt{\underline{D}el} notifications $\delta_s$. We focus on the
last delivery of each \process, since the best is also the last, as
highlighted by Algorithm~\ref{algo:add}.

@@

\begin{definition}[\label{def:lwo}Stale messages]
  Only the latest broadcast of a \node matters.  A message $m$ conveys
  \underline{s}tale control information $\mathcal{S}(m)$ as soon as
  its broadcaster broadcasts another message: $\mathcal{S}(\alpha_x) =
  \exists \delta_x: b_x(\alpha_x) \rightarrow b_x(\delta_x)$.  A
  \process only delivers or sends messages that it assumes up-to-date.
  % : $s_x(\alpha_s)
  % \implies \delta_s \not\in R_x$.
  For convenience, we define
  % \TODO{$\textup{minUTD(}D\textup{)} = \min
  %  \{\alpha_s^d | \alpha_s^d \in D \wedge \delta_s \notin D\}$}, and
  $\textup{last(}D\textup{)} = \alpha_s \in D: \nexists \alpha_{s'}
  \in D: d_x(\alpha_s) \rightarrow d_x(\alpha_{s'})$.
  %  \TODO{Only keep last?  Tothink what happens if we keep all as in
%    def now.}
\end{definition}

A naive attempt at implementing DCP resembles echoes in
acoustics. Following the principles of scoped broadcast as stated in
Definition~\ref{def:scoped}, a \process that receives a staleness
notification forwards it only if the latter targets its current
partition. The staleness notifications propagate through the network
as long as they remain in the deleted partition. When they reach the
bordering \processes of the deleted partition, it creates an echo of
bordering partitions that will go backward to fill the gap left open
by removals using forwarding of best (FB) as stated by
Theorem~\ref{theo:fb}.

@@

\begin{definition}[\label{def:fs}\underline{F}orwarding of
  \underline{s}taleness (FS)]
  %
  Any source can broadcast a staleness notification at any
  time. Every \process $x$ delivers and forwards a received staleness
  notification if it targets its best up-to-date delivered message:
  %% \TODO{$(\delta_s \in R_x \wedge \alpha_s =
  %%     \textup{minUTD}~D_x) \iff \eventually (d_x(\delta_s) \rightarrow
  %% f_x(\delta_s))$.}
  $(\textup{last}~D_x = \alpha_s \wedge d_x(\alpha_s) \rightarrow
  r_x(\delta_s)) \implies d_x(\delta_s) \rightarrow f_x(\delta_s)$.
  % \wedge \nexists d'_s: d_x(\delta_s) \TODO{\not\rightarrow} d_x(\delta_s'))$.
\end{definition}

\begin{definition}[\label{def:fbplus}$\textup{FB}^+$: echos] FB where
  a \process $x$ that receives but does not deliver a staleness
  notification sends back --~or echoes~-- its best up-to-date
  delivered message: $(\textup{last}~D_x = \alpha_s^d \wedge \delta_s
  \not\in D_x \wedge d_x(\alpha_s^d)\rightarrow r_{xy}(\delta_{s'}))
  \implies r_{xy}(\delta_{s'}) \rightarrow
  s_{xy}(\alpha_s^d\oplus^{w_{xy}})$.

  % $r_{xy}(\alpha_s)
  % \wedge \delta_s \in D_x \implies s_{xy}(\alpha_{s'}^d\oplus
  % ^{w_{xy}})$ with $\alpha_{s'^d} = \min D_x \wedge \delta_{s'} \not\in
  % D_x$. \TODO{Rework min}
\end{definition}

@@

\begin{definition}[\label{def:fsplus}$\textup{FS}^+$: forwarding
    staleness downstream] Any \emph{\process} can broadcast a
  staleness notification at any time. A \emph{child \process} $x$
  delivers and forwards a received staleness notification if it comes
  from the path of delivery of its best up-to-date delivered
  message. \TODO{more.}
\end{definition}
